Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC

### 2016-04-02 tese2/11_related_work_collaboration.tex
[ADD] amazing explanation of network test results
[ADD] spell checking, explained private network component at architecture

        %RP lá vais tu mudar de tema da conversa sem explicar porquê....
        Even though one on one calls are common, there are occasions when several people take part of the same video call.
	Multi-party video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants. Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
        In this scenario, a more efficient alternative to peer-to-peer is the use of a Multipoint Control Unit (MCU).

@@

	{\color{blue}
	\ac{KMS} supports transcoding, group communications, recording, mixing, broadcasting, aplying filters, image and sound analysis, for example it allows face and qrcode detection. \ac{KMS} functionalites are exposed through \emph{JSON-RPC} over \emph{WebSocket}, there are three clients available: JavaScript Client for web browsers, Java Client for Java EE Servers and JavaScript Cleint for Node.js servers. 
	\ac{KMS} suports streaming over \emph{WebRTC}, \ac{HTTP} and \ac{RTP} endpoints. Another important component of \ac{KMS} is \emph{Kurento Repository} which suports recording and playing directly from \emph{MongoDB}, that is important for providing a scalable media storage. Unlinke \emph{Jitsi Video Bridge}, \ac{KMS} does not enforce a specific signaling protocol.
	}

@@

	\emph{TogetherJS} is a JavaScript library that uses \ac{WebRTC} for collaborative web applications. It uses \ac{JSON} messages for \ac{OT} concurrency control but it does not provide storage. {\color{blue}\emph{TogetherJS} uses its own servers for the signaling phase and it supports microphone and mouse sharing between users. It requires a simple server (also known as hub) that echoes messages between clients, all the synchronization work is performed by the clients. Altough the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}, if we would not choose \emph{NodeJS} for server implementation we would need read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}

@@

	{\color{blue}\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the text operations but also to associate authorships to pieces of text. It supports adding functionality trought plugins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example:
		\begin{itemize}
		\item exporting and importing document formats such as \emph{DOC}, \emph{PDF}, \emph{ODT} and \emph{DOCX}.
		\item painting and drawing.
		\item video and audio chat using WebRTC.
		\item create slideshows.
		\item spell checking.
		\item text to speech.
		\end{itemize}
	}

@@

\begin{table}[H]
\centering

\label{my-label}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Library} & \textbf{Own Server} & \textbf{Own Storage} & \textbf{Operations} \\ \hline
ShareJS          & yes                 & yes                  & text+objects        \\ \hline
TogetherJS       & yes                 & no                   & text+objects        \\ \hline
Goodow           & yes                 & yes                  & text+objects        \\ \hline
Etherpad Lite    & yes                 & yes                  & extendable 			    \\ \hline
OT.js            & no                  & no                   & text                \\ \hline
\end{tabular}
\caption{Comparision between Operational Transformation libraries}
\end{table}

>>>>

        %RP lá vais tu mudar de tema da conversa sem explicar porquê....
        Even though one on one calls are common, there are occasions when several people take part of the same video call.
	Multi-party video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants. Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
        In this scenario, a more efficient alternative to peer-to-peer is the use of a \ac{MCU}.

@@

	{\color{blue}
	\ac{KMS} supports trans-coding, group communications, recording, mixing, broadcasting, applying filters, image and sound analysis, for example it allows face and qrcode detection. \ac{KMS} functionalities are exposed through \emph{JSON-RPC} over \emph{WebSocket}, there are three clients available: \emph{JavaScript} client for web browsers, \emph{Java} client for \emph{Java EE} servers and \emph{JavaScript} client for \emph{Node.js} servers. 
	\ac{KMS} supports streaming over \emph{WebRTC}, \ac{HTTP} and \ac{RTP} endpoints. Another important component of \ac{KMS} is \emph{Kurento Repository} which supports recording and playing directly from \emph{MongoDB}, that is important for providing a scalable media storage. Unlinke \emph{Jitsi Video Bridge}, \ac{KMS} does not enforce a specific signaling protocol.
	}

@@

	\emph{TogetherJS} is a JavaScript library that uses \ac{WebRTC} for collaborative web applications. It uses \ac{JSON} messages for \ac{OT} concurrency control but it does not provide storage. {\color{blue}\emph{TogetherJS} uses its own servers for the signaling phase and it supports microphone and mouse sharing between users. It requires a simple server (also known as hub) that echoes messages between clients, all the synchronization work is performed by the clients. Although the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}, if we would not choose \emph{NodeJS} for server implementation we would need read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}

@@

	{\color{blue}\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the text operations but also to associate authorships to pieces of text. It supports adding functionality trough plug-ins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example:
		\begin{itemize}
		\item exporting and importing document formats such as \emph{DOC}, \emph{PDF}, \emph{ODT} and \emph{DOCX}.
		\item painting and drawing.
		\item video and audio chat using WebRTC.
		\item create sideshows.
		\item spell checking.
		\item text to speech.
		\end{itemize}
	}

@@

\begin{table}[H]
\centering
\caption{Comparision between Operational Transformation libraries}
\label{my-label}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Library} & \textbf{Own Server} & \textbf{Own Storage} & \textbf{Operations} \\ \hline
ShareJS          & yes                 & yes                  & text+objects        \\ \hline
TogetherJS       & yes                 & no                   & text+objects        \\ \hline
Goodow           & yes                 & yes                  & text+objects        \\ \hline
Etherpad Lite    & yes                 & yes                  & extendable 			    \\ \hline
OT.js            & no                  & no                   & text                \\ \hline
\end{tabular}
\end{table}

### 2016-04-02 tese2/13_implementation_model.tex
[ADD] amazing explanation of network test results
[ADD] performance tests, updated playFW, websocket simplification
[ADD] spell checking, explained private network component at architecture

\section{Model}
The model is the most important component of our solution, a badly designed model can imply serious deficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.

@@

Our first decision for our model in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes indepently from the object type. To this generic set of properties we call data and each object of this type has a reference to the owner, which is a unique identification number.

\begin{table}[H]
\centering
\label{my-label}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Data}}                             \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ owner           & ObjectId                         \\ 
            $\medbullet$ properties      & Document                         \\ 
            $\medcirc$ searchableValue & List\textless Text\textgreater     \\ 
            $\medcirc$ identiableKeys  & List\textless Text\textgreater     \\ \hline
        \end{tabular}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{DataPermission}}                   \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ owner           & ObjectId                         \\ 
            $\medcirc$ readPermissions & Document                           \\ 
            $\medcirc$ writePermissions  & Document                         \\ \hline
        \end{tabular}    
    \end{tabular}
\caption{Generic data model}
\end{table}

@@

Another important attribute specification is the owner identifiability, which tells us if the attribute identifies the object. This specification let us create abstract authentication services, for example a user can login into our system by providing any attribute that identifies himself, for example the e-mail but others are possible like the username or cellphone number.

@@

The user model is not tied to the user attributes, the information maintained in this model is just used for authentication purposes. Passwords are not stored in plaintext, instead we apply hashing and salting techniques \cite{password} in order to make it harder to decode the password by an attacker. We use \emph{SHA-1} and a random salt per user with 32 characters long.

\begin{table}[H]
\centering
\label{my-label}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{User}}                             \\ \hline
        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
        $\medbullet$ hash           & Text                         \\ 
        $\medbullet$ salt      & Text                         \\ \hline
    \end{tabular}
\caption{User model}
\end{table}

@@

\begin{table}[H]
\centering
\label{my-label}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Relation}}                             \\ \hline
        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
        $\medbullet$ source           & ObjectId                         \\ 
        $\medbullet$ target      & ObjectId                         \\ \hline
    \end{tabular}
\caption{Relation model}
\end{table}

@@

A group can be public or private. If the group is public then it is visible to all users that mantain a friendship with a member of this group. If the group is private then it is only visible to their members.

\begin{table}[H]
\centering
\label{my-label}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Group}}                             \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medcirc$ inviteToken           & Text                         \\ 
            $\medbullet$ visibility      & Text                         \\ \hline
        \end{tabular}

        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{GroupMembership}}                   \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ groupId           & ObjectId                         \\ 
            $\medbullet$ userId & ObjectId                           \\ \hline
        \end{tabular}    
    \end{tabular}
\caption{Group model}
\end{table}

@@

\begin{table}[H]
\centering
\label{my-label}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Message}}                             \\ \hline
        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
        $\medbullet$ source           & ObjectId                         \\ 
        $\medbullet$ target      & ObjectId                         \\ 
        $\medbullet$ sequence      & Number                         \\ 
        $\medbullet$ content      & Text                         \\ \hline
    \end{tabular}
\caption{Message model}
\end{table}

@@

\begin{table}[H]
\centering
\label{my-label}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{HyperContent}}                             \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ groupId            & ObjectId                         \\ 
            $\medbullet$ start              & Date                         \\ 
            $\medbullet$ end                & Date                         \\ 
            $\medbullet$ content            & Text                         \\ \hline
        \end{tabular}

@@

        \end{tabular}    
    \end{tabular}
    \caption{Hyper content model}
\end{table}

@@

\begin{table}[H]
\centering
\label{my-label}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{CollaborativeContent}}                             \\ \hline
        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
        $\medbullet$ groupId            & ObjectId                         \\ 
        $\medbullet$ content            & Text                         \\ \hline
    \end{tabular}
    \caption{Collaborative content model}
\end{table}

@@

During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks. Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains a reference to a group and a mapping between entity that was recorded and the correspondent video url.

\begin{table}[H]
\centering
\label{my-label}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{RecordingInterval}}                   \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ groupId            & ObjectId                         \\ 
             $\medbullet$ start             & Date                         \\ 
            $\medbullet$ end                & Date                         \\ \hline
        \end{tabular}   
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{RecordingChunk}}                             \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ groupId            & ObjectId                         \\ 
            $\medbullet$ interval           & ObjectId                      \\ 
            $\medbullet$ start              & Date                         \\ 
            $\medbullet$ end                & Date                         \\ 
            $\medbullet$ sequenceNumber     & Number                         \\ 
            $\medbullet$ urls               & Document                       \\ \hline
        \end{tabular}
    \end{tabular}
    \caption{Recording model}
\end{table}

>>>>

\section{Model}
The model is the most important component of our solution, a badly designed model can imply serious difficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.

@@

Our first decision for our model in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes independently from the object type. To this generic set of properties we call data and each object of this type has a reference to the owner, which is a unique identification number.

\begin{table}[!htb]
\centering
\caption{Generic data model}
\label{my-label}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Data}}                             \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ owner           & ObjectId                         \\ 
            $\medbullet$ properties      & Document                         \\ 
            $\medcirc$ searchableValue & List\textless Text\textgreater     \\ 
            $\medcirc$ identiableKeys  & List\textless Text\textgreater     \\ \hline
        \end{tabular}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{DataPermission}}                   \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ owner           & ObjectId                         \\ 
            $\medcirc$ readPermissions & Document                           \\ 
            $\medcirc$ writePermissions  & Document                         \\ \hline
        \end{tabular}    
    \end{tabular}
\end{table}

@@

Another important attribute specification is the owner identifiability, which tells us if the attribute identifies the object. This specification let us create abstract authentication services, for example a user can login into our system by providing any attribute that identifies himself, for example the e-mail but others are possible like the user name or cellphone number.

@@

The user model is not tied to the user attributes, the information maintained in this model is just used for authentication purposes. Passwords are not stored in plain text, instead we apply hashing and salting techniques \cite{password} in order to make it harder to decode the password by an attacker. We use \emph{SHA-1} and a random salt per user with 32 characters long.

\begin{table}[!htb]
\centering
\caption{User model}
\label{my-label}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{User}}                             \\ \hline
        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
        $\medbullet$ hash           & Text                         \\ 
        $\medbullet$ salt      & Text                         \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
\caption{Relation model}
\label{my-label}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Relation}}                             \\ \hline
        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
        $\medbullet$ source           & ObjectId                         \\ 
        $\medbullet$ target      & ObjectId                         \\ \hline
    \end{tabular}
\end{table}

@@

A group can be public or private. If the group is public then it is visible to all users that maintain a friendship with a member of this group. If the group is private then it is only visible to their members.

\begin{table}[!htb]
\centering
\caption{Group model}
\label{my-label}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{Group}}                             \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medcirc$ inviteToken           & Text                         \\ 
            $\medbullet$ visibility      & Text                         \\ \hline
        \end{tabular}

        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{GroupMembership}}                   \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ groupId           & ObjectId                         \\ 
            $\medbullet$ userId & ObjectId                           \\ \hline
        \end{tabular}    
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
\caption{Message model}
\label{my-label}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{Message}}                             \\ \hline
        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
        $\medbullet$ source           & ObjectId                         \\ 
        $\medbullet$ target      & ObjectId                         \\ 
        $\medbullet$ sequence      & Number                         \\ 
        $\medbullet$ content      & Text                         \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
\caption{Hyper content model}
\label{my-label}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{HyperContent}}                             \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ groupId            & ObjectId                         \\ 
            $\medbullet$ start              & Date                         \\ 
            $\medbullet$ end                & Date                         \\ 
            $\medbullet$ content            & Text                         \\ \hline
        \end{tabular}

@@

        \end{tabular}    
    \end{tabular}
\end{table}

@@

\begin{table}[!htb]
\centering
    \caption{Collaborative content model}
    \label{my-label}
    \begin{tabular}{|ll|}
        \hline
        \multicolumn{2}{|c|}{\textbf{CollaborativeContent}}                             \\ \hline
        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
        $\medbullet$ groupId            & ObjectId                         \\ 
        $\medbullet$ content            & Text                         \\ \hline
    \end{tabular}
\end{table}

@@

During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks. Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains a reference to a group and a mapping between entity that was recorded and the correspondent video \ac{URL}.

\begin{table}[!htb]
\centering
    \caption{Recording model}
    \label{my-label}
    \begin{tabular}{cc}
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{RecordingInterval}}                   \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ groupId            & ObjectId                         \\ 
             $\medbullet$ start             & Date                         \\ 
            $\medbullet$ end                & Date                         \\ \hline
        \end{tabular}   
        \begin{tabular}{|ll|}
            \hline
            \multicolumn{2}{|c|}{\textbf{RecordingChunk}}                             \\ \hline
            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
            $\medbullet$ groupId            & ObjectId                         \\ 
            $\medbullet$ interval           & ObjectId                      \\ 
            $\medbullet$ start              & Date                         \\ 
            $\medbullet$ end                & Date                         \\ 
            $\medbullet$ sequenceNumber     & Number                         \\ 
            $\medbullet$ urls               & Document                       \\ \hline
        \end{tabular}
    \end{tabular}
\end{table}

### 2016-04-02 tese2/14_implementation_signaling.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)
[ADD] spell checking, explained private network component at architecture

The signaling protocol could be implemented using \ac{HTTP} messages but they would transport extra information such as \ac{HTTP} headers and would follow a request response signaling mechnism which would not be the best option, as multiple ice candidates can arrive at any time. Using a \ac{REST} \ac{API} over \ac{HTTP} for signaling and allowing other methods to be called at the same time would lead to open multiple \ac{TCP} connections at the same time.

Instead if we create a \emph{WebSocket} \ac{API} we only need one \ac{TCP} connection and at the same provide bi-directional communications without additional headers.

@@

When a user enters in a group conference, after the page is completely loaded it is created a WebSocket that mantains a connection with our web servers. 
Before creating the web socket, we must identify the user and check if he has permissions to participate in the conference. The user identification is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers then we retrieve all the information need from database in order to check if the user has permissions to join the conference room. It's important to save de user identification before the socket is created because after the the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455} the \ac{HTTP} context is lost.

When the connection is established between the server and the client, a \emph{PeerConnection} is created on the client and imediately after an \ac{WebRTC} endpoint on the server specifying a possible set of \ac{ICE} servers to connect.

The user is asked if he whants to share its camera and microphone, share screen or just receive streams from the server. If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2015).} may ask to install a pluggin if the browser doesn't support screen sharing.

If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent contraints in order to obtain a local stream. Use the constraints presented on listing \ref{lst:constraints}.

@@

A popup is raised in order to ask the user to give permission to share those resources. If the resources are shared successfully the user creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets. If the resources are not shared or the user specified to receive stream only, an offer is created specifying the contraints for receiving only video and audio.

@@

The \emph{local session description} contains the session identifier, codecs, containers, transport protocols and ports used per media type. The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record and analyse streams coming from the user \ac{WebRTC} endpoint.

The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint. Then a \emph{local session description} like listing \ref{lst:sig02} is created on the server and sent back to the client, after that the server trys to gather \ac{ICE} candidates.

@@

After a while both the server and client receive the \ac{ICE} candidates that allow the client to connect directly to \ac{KMS} and vice-versa. The candidates are received at the client which sets them to its \emph{PeerConnection}. The same is done on the server which receives the \ac{ICE} candidates like on listing \ref{lst:sig03} from the client and propagates them to \ac{KMS}.

@@

Both intervenients test the conectivity of each \ac{ICE} candidate. When a connection is established the user and server start to interchange stream data but other \ac{ICE} candidates may arrive with better connections, when that happens the connection changes seamlessly.

With the media session established, the server starts to record any received stream and the client creates an \ac{URL} correspondent to the stream location.

\begin{figure}[!htb]
    \centering
    \begin{subfigure}{}
    	\includegraphics[width=0.8\textwidth]{figures/signaling}
    \end{subfigure}
    \caption{Signaling sequence diagram}
\end{figure}

>>>>

The signaling protocol plays a fundamental role on our solution, after the web server validates the user access,
it allows the users to directly connect to \ac{KMS} which is placed in a private network and let the server and users negotiate media types and encoding information to use during the conversation.

With respect to the signaling protocol, it could be implemented using \ac{HTTP} messages but they would transport extra information such as \ac{HTTP} headers and would follow a request-response signaling mechanism which would not be the best option, as multiple ice candidates can arrive at any time. Using a \ac{REST} \ac{API} over \ac{HTTP} for signaling and allowing other methods to be called at the same time would lead to open multiple \ac{TCP} connections at the same time.

Instead if we create a \emph{WebSocket} \ac{API} we only need one \ac{TCP} connection and at the same provide bi-directional communications without additional headers. Moreover, using \emph{WebSocket} allows the server to send messages to the client without the client having to request it, which can be slightly faster than using \ac{HTTP}.

@@

When a user enters in a group conference, after the page is completely loaded it is created a WebSocket that maintains a connection with our web servers. 
Before creating the web socket, we must identify the user and check if he has permissions to participate in the conference. The user identification is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers then we retrieve all the information need from database in order to check if the user has permissions to join the conference room. It's important to save the user identification before the socket is created because after the the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455} the \ac{HTTP} context is lost.

When the connection is established between the server and the client, a \emph{PeerConnection} is created on the client and immediately after an \ac{WebRTC} endpoint on the server specifying a possible set of \ac{ICE} servers to connect.

The user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server. If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2015).} may ask to install a plug-in if the browser doesn't support screen sharing.

If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent constraints in order to obtain a local stream. Use the constraints presented on listing \ref{lst:constraints}.

@@

A pop-up is raised in order to ask the user to give permission to share those resources. If the resources are shared successfully the user creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets. If the resources are not shared or the user specified to receive stream only, an offer is created specifying the constraints for receiving only video and audio.

@@

The \emph{local session description} contains the session identifier, codecs, containers, transport protocols and ports used per media type. The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record and analyze streams coming from the user \ac{WebRTC} endpoint.

The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint. Then a \emph{local session description} like listing \ref{lst:sig02} is created on the server and sent back to the client, after that the server tries to gather \ac{ICE} candidates.

@@

Subsequently, after a while both the server and client receive the \ac{ICE} candidates that allow the client to connect directly to \ac{KMS} and vice-versa. The candidates are received at the client which sets them to its \emph{PeerConnection}. The same is done on the server which receives the \ac{ICE} candidates like on listing \ref{lst:sig03} from the client and propagates them to \ac{KMS}.

@@

In addition both intervenients test the connectivity of each \ac{ICE} candidate. When a connection is established the user and server start to interchange stream data but other \ac{ICE} candidates may arrive with better connections, when that happens the connection changes seamlessly.

Having the media session established, the server starts to record any received stream and the client creates an \ac{URL} correspondent to the stream location.

\begin{figure}[!htb]
    \centering
    \begin{subfigure}{}
    	\includegraphics[width=0.9\textwidth]{figures/signaling}
    \end{subfigure}
    \caption{Signaling sequence diagram}
\end{figure}

### 2016-04-02 tese2/15_implementation_recording.tex
[ADD] change architecture
[ADD] spell checking, explained private network component at architecture

	After recording, each block was uploaded to our server through \ac{HTTP} and saved into the filesystem, the block metadata was created and inserted on our database. Each block metadata contained the file's location for the recorded block, the starting date, duration, user identification and group identification. When the file was completely saved the metadata was advertised to the remaining users. This metadata was simply used to refresh the user interface, the metadata was completely discard after that because an huge amount of small blocks metadata would use more and more memory as time went by.

@@

	Altough this idea was fairly simple, we couldn't achieve seamless sequential block switching. Downloading the video file always toke a noticeable amount of time. To solve this problem while we were playing a block, the next one was downloading in parallel so when the current block finished playing we would have an available block to play.

	Block switching became more acceptable, but switching the url always produced a flash. We solved this problem by having two layers and set the url to the back layer, the front layer freezed with the last frame, then we changed the back layer to front.

	After we implemented our recording solution using this approach, we tested localy and remotely. For remote connections we observed a fairly high bandwidth usage mainly because blocks were both sent and received at maximum quality.

\subsection{Server side recording to filesystem}

	Before recording any type of stream we had to analyse the user media offer in order to check if a video as realy being received by \ac{KMS}, otherwise if we would not verify the user's offer, the recording video would be black.

	The streaming content received on \ac{KMS} was already compressed due to \ac{WebRTC}'s exchanged \ac{QoS} metrics data. As a direct consequence our recorder solution used on most cases less disk space per block, but would never use more storage than client side recording. \ac{KMS} allows recording files using \emph{webm} and \emph{mp4} containers.

	With server side recording, the user would maintain always the same stream url even if it is playing realtime video or reproducing recorded video. When a user desires to play recorded video, a websocket message is sent specifying the time and the intended user id, the group identification is not sent because it is already associated to the websocket. The server performs the same calculations in order to find a block that intersects the requested time, plays it and when finished the next part is automatically played wihtout the user intervention.

@@

	Back when we were implementing our solution, \ac{KMS} had not support for seeking videos, which meant that blocks would would always start playing from their beggining. This lead to a theoric playing time error that can be at most half the duration of a block. In practice the maximum playing time error coincides with the block duration because we perform the intersection between the requested date and the block but we could decrease the error by half if we calculate the intersection with the requested time plus half the duration.

@@

	\end{tikzpicture}
		\caption{Theoretic maximum playing error}
	\end{figure}

@@

	\end{tikzpicture}
		\caption{Practical maximum playing error}
	\end{figure}

	For playing video with an higher velocity we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/}(accessed: 17 March 2016)} to convert the block into a new video with the desired velocity and seek time. Because the media duration is known, when the video started to being convert the headers located at the begining of the file were already written and that made it possible to stream while converting.

@@

	Although we implemented a solution that worked, we imediately noticed that \emph{ffmpeg} would take some time to initialize and that lead to pauses between switching parts.

@@

	One of our concerns during the development of our solution was the storage scalability. Saving files directly into filesystem would require an extra effort to distribute and replicate files among servers. For that purpose \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.

	One of the features that \emph{Kurento Repository} provides is the ability to play directly from the database without having to download the entire file to \ac{KMS}. The same is true for recording, but because the file headers are in the beggining and the file is written until is stops, the headers don't contain the necessary information for seeking the file.

>>>>

	After recording, each block was uploaded to our server through \ac{HTTP} and saved into the file system, the block meta-data was created and inserted on our database. Each block meta-data contained the file's location for the recorded block, the starting date, duration, user identification and group identification. When the file was completely saved the meta-data was advertised to the remaining users. This meta-data was simply used to refresh the user interface, the meta-data was completely discard after that because an huge amount of small blocks meta-data would use more and more memory as time went by.

@@

	Although this idea was fairly simple, we couldn't achieve seamless sequential block switching. Downloading the video file always toke a noticeable amount of time. To solve this problem while we were playing a block, the next one was downloading in parallel so when the current block finished playing we would have an available block to play.

	Block switching became more acceptable, but switching the \ac{URL} always produced a flash. We solved this problem by having two layers and set the \ac{URL} to the back layer, the front layer frozen with the last frame, then we changed the back layer to front.

	After we implemented our recording solution using this approach, we tested locally and remotely. For remote connections we observed a fairly high bandwidth usage mainly because blocks were both sent and received at maximum quality.

\subsection{Server side recording to file system}

	Before recording any type of stream we had to analyze the user media offer in order to check if a video as really being received by \ac{KMS}, otherwise if we would not verify the user's offer, the recording video would be black.

	The streaming content received on \ac{KMS} was already compressed due to \ac{WebRTC}'s exchanged quality of service metrics data. As a direct consequence our recorder solution used on most cases less disk space per block, but would never use more storage than client side recording. \ac{KMS} allows recording files using \emph{webm} and \emph{mp4} containers.

	With server side recording, the user would maintain always the same stream \ac{URL} even if it is playing real-time video or reproducing recorded video. When a user desires to play recorded video, a \emph{webSocket} message is sent specifying the time and the intended user id, the group identification is not sent because it is already associated to the \emph{webSocket}. The server performs the same calculations in order to find a block that intersects the requested time, plays it and when finished the next part is automatically played without the user intervention.

@@

	Back when we were implementing our solution, \ac{KMS} had not support for seeking videos, which meant that blocks would would always start playing from their beginning. This lead to a theoretic playing time error that can be at most half the duration of a block. In practice the maximum playing time error coincides with the block duration because we perform the intersection between the requested date and the block but we could decrease the error by half if we calculate the intersection with the requested time plus half the duration.

@@

	\end{tikzpicture}
		\caption{Theoretic maximum playing error $(e)$}
	\end{figure}

@@

	\end{tikzpicture}
		\caption{Practical maximum playing error $(e)$}
	\end{figure}

	For playing video with an higher velocity we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/}(accessed: 17 March 2016)} to convert the block into a new video with the desired velocity and seek time. Because the media duration is known, when the video started to being convert the headers located at the beginning of the file were already written and that made it possible to stream while converting.

@@

	Although we implemented a solution that worked, we immediately noticed that \emph{ffmpeg} would take some time to initialize and that lead to pauses between switching parts.

@@

	One of our concerns during the development of our solution was the storage scalability. Saving files directly into file system would require an extra effort to distribute and replicate files among servers. For that purpose \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.

	One of the features that \emph{Kurento Repository} provides is the ability to play directly from the database without having to download the entire file to \ac{KMS}. The same is true for recording, but because the file headers are in the beginning and the file is written until is stops, the headers don't contain the necessary information for seeking the file.
